Spherical harmonic analysis of faceted spheroids identifies shaping strategies and standardisation at Qianshangying (North China)

### 2025-08-22 analysis/paper/paper.qmd
editing

Additionally, we computed the spherical harmonic energy (SHE), defined as the sum of squared normalized spherical harmonic coefficients for each degree. This provides a rotation-invariant measure of shape complexity across spatial frequencies. For normalized shapes, this total energy approaches one for a perfect sphere, while more irregular forms yield higher values.

Notably, classic sphericity metrics [@Wadell1935] exhibit critical limitations in differentiating morphotypes with equivalent surface-area-to-volume ratios. SHE overcomes this constraint through spherical harmonic decomposition of surface morphology, and this allows it to distinguish localized irregularities (like core edge or concave) from global shape deviations (e.g., elongation) – differences that classic sphericity metrics often fail to capture.

@@

We applied diacritical analysis to reconstruct the reduction sequence based on superimposition relationships among scars on spheroids [@tittonSubspheroidsLithicAssemblage2020; @Cabanes2024]. This method enables the identification of scars order, allowing us to figure out the organization of removals. All spheroids, polyhedrons and multifacial cores are included in the analysis to assess whether removals organization differs systematically within these groups.

>>>>

Additionally, we computed the spherical harmonic energy (SHE), defined as the sum of squared normalized spherical harmonic coefficients for each degree. This provides a rotation-invariant measure of shape complexity across spatial frequencies. For normalized shapes, this total energy approaches a value of one for a perfect sphere, while more irregular forms yield higher values.

Notably, classic sphericity metrics [@Wadell1935] exhibit critical limitations in differentiating morphotypes with equivalent surface-area-to-volume ratios. SHE overcomes this constraint through spherical harmonic decomposition of surface morphology, and this allows it to distinguish localized irregularities (like core edge or concave) from global shape deviations (e.g., elongation)--differences that classic sphericity metrics often fail to capture.

@@

We used diacritical analysis to reconstruct the reduction sequence based on superimposition relationships among scars on spheroids [@tittonSubspheroidsLithicAssemblage2020; @Cabanes2024]. This method enables the identification of the order of scar removal, allowing us to understand the organization of removals. All spheroids, polyhedrons and multifacial cores were included in the analysis to assess whether removal organization differs systematically within these groups.
